\section{Lattice based authenticated key-exchange}
Key-exchange protocol (KEX) is a cryptographic primitive to derive a common secret key via a public network communication between a number of parties. The parties do not share any secret information beforehand. KEX that also authenticates the identities of the involved parties is called authenticated key-exchange protocol (AKE). More formally, in an AKE protocol each party has a static public-secret key pair. The static public key is certified with the party's identity. When running the protocol, each party generates an ephemeral secret key and, depending on that, a corresponding ephemeral public key. The public key is sent to the other party. Each party computes a shared session key by using the ephemeral key and the static key (e.g. password).

An authenticated key-exchange protocol is said to have perfect forward secrecy (PFS) if a compromise of its static keys does not lead to a compromise of previously established and deleted session keys. There are many existing authenticated key-exchange approaches built for Diffie-Hellman, however they all can be used with Ring-$\mathrm{LWE}$ instead because of similarities in design. Authenticated key-exchange aims to overcome man-in-the-middle attack, which is an attack where the attacker secretly relays and possibly alters the communication between two parties who believe they are directly communicating with each other. The following are basic authenticated Diffie-Hellman approaches that can be applied to RLWE-KEX:

\begin{enumerate}
    \item Use key-exchange in conjunction with a signing certificate (use either RSA or DSA as a public key). The server sends its certificate which client validates and ephemeral key signed (or encrypted) using server's private key. Client validates the ephemeral key using certificate and encrypts the session using ephemeral key. 
    \item Concatenate (or pad) ephemeral key (or result of unauthenticated key-exchange) with a password and use the hash of that as a session key to achieve a secure communication
    \item Encrypt $b = A s + e$ (in Ring-$\mathrm{LWE}$) or $K = g^a \bmod p$ (in Diffie-Hellman) using a password as an encryption key before sending it to other party.
\end{enumerate}